---
title: "DataFest2020_ExplorationAnalysis_AJ"
author: "Ashlyn Jew"
date: "5/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load data
```{r}
crime2020 <- read.csv("Crime_Data_from_2020_to_Present.csv")
crime2010 <- read.csv("Crime_Data_from_2010_to_2019.csv")
crime <- rbind(crime2010, crime2020)
```

#Clean data
```{r}
library(tidyverse)

#delete DR_NO, Rpt.Dist.No, Part.1-2, Mocodes, Crm.Cd.1, Crm.Cd.2, Crm.Cd.3, Crm.Cd.4 columns
crime <- crime[-c(1, 7, 8, 11, 21, 22, 23, 24)]

#rename columns
names(crime) <- c("Date_Rptd", "Date_Occ", "Time_Occ", "Area_Code", "Area_Name", "Crm_Code", "Crm_Desc", "Vict_Age", "Vict_Sex", "Vict_Descent", "Premis_Code", "Premis_Desc", "Weapon_Code", "Weapon_Desc", "Status_Code", "Stat_Desc", "Location", "Cross_Street", "Lat", "Lon")

#convert Date.Rptd, DATE.OCC from factor to Date
crime$Date_Rptd <-  as.Date(crime$Date_Rptd, format = "%m/%d/%Y")
crime$Date_Occ <- as.Date(crime$Date_Occ, format = "%m/%d/%Y")

#filter for just 2019 and 2020 data
crime <- crime %>% filter(Date_Occ > "2018-12-31")
#arrange by ascending date and time
crime <- crime %>% arrange(Date_Occ, Time_Occ)
```

#Exploration Analysis
```{r}
library(ggplot2)

crime_1920 <- crime %>% filter(DATE.OCC > "2018-12-31")
crime_20 <- crime %>% filter(DATE.OCC > "2019-12-31")

ggplot(data = crime_20,
       mapping = aes(AREA.NAME,
                     fill = factor(month(crime_20$DATE.OCC)))) + geom_histogram(stat = "count")
```

Doesn't seem like there's any significant increase in crime.

```{r}
table(crime_20$DATE.OCC)
plot(table(crime_20$DATE.OCC), type = "l")
```

```{r}
crime_20 %>% 
  filter(Crm.Cd.Desc %in% c(c("BURGLARY", "ROBBERY"))) %>% 
  group_by(month(DATE.OCC)) %>% 
  summarise(
    n = n()
  )

crime_1920 %>% 
  filter(Crm.Cd.Desc %in% c(c("BURGLARY", "ROBBERY"))) %>% 
  group_by(year(DATE.OCC), month(DATE.OCC)) %>% 
  summarise(
    n = n()
  )
```

Seems like there's like a <100 decrease in burglary/robbery each month compared to last year

```{r}
crime_1920 %>% 
  group_by(year(DATE.OCC), month(DATE.OCC)) %>% 
  summarise(
    n = n()
  )
```
In March and April there's a 3000ish decrease in overall reported crimes

```{r}
crime_20 %>% 
  group_by(month(DATE.OCC), AREA.NAME) %>% 
  summarise(
    n = n()
  )
```

```{r}

#Descent Code: A - Other Asian B - Black C - Chinese D - Cambodian F - Filipino G - Guamanian H - Hispanic/Latin/Mexican I - American Indian/Alaskan Native J - Japanese K - Korean L - Laotian O - Other P - Pacific Islander S - Samoan U - Hawaiian V - Vietnamese W - White X - Unknown Z - Asian Indian
crime_20 %>% 
  group_by(month(DATE.OCC), Vict.Descent) %>% 
  summarise(
    n = n()
  )

crime_1920 %>% 
  filter(Vict.Descent == "C") %>% 
  group_by(year(DATE.OCC), month(DATE.OCC), Vict.Descent) %>% 
  summarise(
    n = n()
  )
```

Doesn't seem like there's any increase in Chinese victims compared to last year.